\documentclass[11pt]{article}
\usepackage{amsmath}
\usepackage{bussproofs}
\usepackage{caption}
\usepackage[margin=4cm]{geometry}
\usepackage[hyphens]{url}
\usepackage{listings}
\usepackage{titling}
\usepackage{xcolor}

% Code listings
\lstset{basicstyle=\small\sffamily,
        breaklines=true,
        language=Haskell,
        columns=flexible,
        keepspaces=true,
        tabsize=4,
        commentstyle=\color{gray},
        morekeywords={throwError},
        literate= {<-}{$\leftarrow$}{1}{->}{$\rightarrow$}{1},
        deletekeywords={fst,snd,zero}
      }

% inline proof trees
\newenvironment{bprooftree}
    {\leavevmode\hbox\bgroup}
    {\DisplayProof\egroup}

% Figure titles
\captionsetup{labelfont=bf,font=bf,skip=0pt}

\title{\bf Implementation of Hindley-Milner \\ Type Inference}
\author{Tien Thuy Ho}
\predate{}
\postdate{}
\date{}

\begin{document}
\maketitle

\begin{abstract}
Hindley-Milner type inference system is known for inferring the most general type of a program 
without the use of type annotations. This paper details an implementation of Hindley-Milner type inference 
using the linear-time algorithm W.
\end{abstract}

\section{Introduction} \input{Introduction}
\section{The $\lambda$-Calculus} \input{Language}
\section{Hindley-Milner Type Inference} \input{TypeInference}
\section{Implementation} \input{Implementation}
\section{Conclusion} \input{Conclusion}

\appendix
\input{Appendix}

\begin{thebibliography}{9}
    \bibitem{damas} 
    L. Damas and R. Milner. 
    Principle types-schemes for functional programs. 
    In \textit{POPL ’82: Proceedings of the 9th ACM SIGPLAN-SIGACT symposium on Principles of programming languages}, 
    pages 207-212, 1982.

    \bibitem{grant} 
    I. Grant.
    The Hindley-Milner Type Inference Algorithm.
    January 2011.

    \bibitem{heeran} 
    B. Heeran, J. Hage, and D. Swierstra.
    Generalizing Hindley-Milner Type Inference Algorithms.
    July 2002.

    \bibitem{hutton} 
    Graham Hutton. 
    \textit{Programming in Haskell}. 
    University Press, Cambridge, 2007.

    \bibitem{pierce} 
    Benjamin C. Pierce. 
    \textit{Types and Programming Languages}. 
    The MIT Press, Cambridge, 2002.

    \bibitem{source} 
    Source code, \\
    \url{https://github.com/thuytien140894/AlgorithmW}
\end{thebibliography}
\end{document}